The Array of Long Baseline Antennas for Taking Radio Observations from the Sub-Antarctic

### 2020-10-30 response_to_referee.tex

\noindent \textcolor{blue}{These sections seem fine, but some more
  information about the LWA antenna below 10 MHz would be useful. The
  sentence ``LWA antenna is not optimized for observations below 10
  MHz...'' seems an understatement given that $<10$ MHz is outside the
  published frequency capability of the antenna. A more informed
  statement about sensitivity (or lack thereof would be useful). For
  example, the lack of Galactic Plane and RFI signal below 8 MHz in
  various plots (eg Figs 8, 11) looks more like there might be
  complete loss of sensitivity. Please clarify.}

We have included a new figure (9) that illustrates the simulated $S11$
for the antenna, in addition to an overplot of the median autospectrum
with a crude sky signal estimate given by $(1-S11)$ multiplied with a
$\nu^{-2.6}$ spectrum.

\noindent \textcolor{blue}{The review of 2-level correlator efficiency
  vs S/N ratio is interesting and a nice reminder to radio astronomers
  about the very different regime such a telescope could be
  operating. But the last paragraph of Section 4.2 dismisses the
  suggestion that the S/N ratio could be non-zero with a weird
  argument about Gaussian random fields. But the reality is if there
  is a massive continuum source in the field of view (Crab Nebula,
  Sun, Galactic Plane which all have structure on size scales $<1$
  deg), then this might indeed be a huge concern. The radio continuum
  sky is obviously not a Gaussian random field.}

We agree that the sky is certainly not a Gaussian random field, but to
get well away from the SNR$<$1 regime requires that a compact source
dominate the total power coming into the antenna, not just the power
in the visibility.  For most radio telescopes this would absolutely be
a concern, but the primary beam of ALBATROS is huge.  We already know
from e.g.\ EDGES that the total power is dominated by the Galactic
signal, and in PRIZM, the Crab (and typically the sun) are not obvious
in total power.  Due to ionospheric effects ALBATROS will not
generally be conducting daytime observing, and using the GSM and Crab
fluxes in the 10-80 MHz range, we expect the Crab is only a $\sim10$\%
effect in the total power, as long as the Galaxy is not self-absorbed.
Similarly, in Fig. 8, for a 100m baseline, the visibilities seen by
ALBATROS are only a few percent of the total power.  The text in
section 4.2 has been updated to include these details.

### 2020-10-30 response_to_referee.tex

\noindent \textcolor{blue}{3) Further on the 1-bit quantisation
  approach, would you not only need a better matched antenna but also
  an equalised analogue system to remove the extremely high slope from
  the synchrotron emission at low frequencies. This can be up to 20dB
  at these frequencies. Furthermore, despite talking about 1-bit
  quantisation, as far as I know the plots in the paper are all using
  4 or more bit? Has this been tested in the field or in the lab. I do
  not doubt it will work but it will need careful levelling and
  testing and ideally should be coupled with a noise injection signal
  for calibration.}

In response to both comments above, the SNAP processes the baseband
data using the full resolution of the ADCs (8-bit, with a 14-bit
version possibly in the future) and internally works at 18 bits.  It
is only after channelization that the data are downsampled, so
happily, concerns about cross-channel leakage (either due to RFI,
slopes across the band, antenna performance etc.) aren't affected by
the bit depth at which we save the data.  We have now clarified this
point in section 4.1 of the text and thank the referee for catching
the omission.

\noindent \textcolor{blue}{4) Regarding calibration of this
  instrument, there is virtually no mention of it. This is supposed to
  be a precision cosmology instrument and therefore I would have
  expected some more discussion on calibration. Even if not going down
  the route of an EDGES-style approach, having e.g. a Dicke switch in
  there would allow some measure of the scale and offset.}

First, we stress ALBATROS is not meant to be a precision cosmology
instrument---that will have to wait for future telescope generations.
Dicke switching, at least as part of regular operations, won't add
very much because we can calibrate gain variations off the Galactic
signal.  Each station saves total power autospectra, which have
enormous SNR and are repeated day-to-day.  We find the instrument is
very stable, with RMS gain fluctuations \<0.6\% over several days,
and typical gain calibration errors of \<0.04\% on 3-second timescales
based on the galactic signal.  Any noise offsets will in general not
contribute to cross-correlations and so the chief concern is to remove
them when developing a galactic model.  PRIZM, which has several
levels of calibration switching, is co-located with ALBATROS, has
similar beams, and operates in a similar frequency range (\>30 MHz).
We can derive an approximate absolute calibration by comparing the
diurnal variation in galactic signal (which is insensitive to any
offsets) bewteen PRIZM and ALBATROS.  We have updated the text to
reflect these numbers.

### 2020-10-30 albatros_paper.tex
added more responses on calibration

To assess the loss of sensitivity as the signal level increases, we
use \autoref{eqn:1bit_output} to derive the correlator efficiency
$\epsilon$, which is the ratio of measured signal-to-noise ratio (SNR)
to the ideal, infinite-precision SNR.  \autoref{fig:1bit_efficiency}
illustrates that the decrease in $\epsilon$ as a function of SNR is
relatively shallow.  When the SNR is near zero, $\epsilon$ approaches
$2/\pi \approx 0.64$, and if the signal level increases to two times
the noise level, $\epsilon$ decreases by only a factor of $\sim2$.
The final \albatros\ array will have baselines that are long relative
to the observing wavelength, and the correlated component of the
electric field (which sets the relevant signal level) is likely a
small fraction of the total received power.  This qualitative argument
can by made by considering antennas in the $uv$~plane in the flat-sky
approximation.  Any individual antenna receives all power present in
the $uv$~plane (plus system noise), but the correlated contribution
comes only from the portion of the $uv$~plane that is encompassed by
the primary beam at the antenna's $uv$-space coordinate.  Assuming
that the sky can be described as a Gaussian random field, the maximum
correlated signal between two antennas is determined by the aperture
filling factor of that pair.  For the case of dipole antennas, the
filling factor is the square of the wavelength divided by the baseline
length.  Therefore, in the long-baseline limit, the correlated signal
is small in comparison to the total power received, and the correlator
efficiency is unlikely to decrease significantly from the ideal
$2/\pi$ value.  This efficiency behavior remains true even when the
brightest radio sources are visible, since the primary beam
encompasses more than a steradian solid angle (the sun is irrelevant
since ionospheric effects restrict \albatros\ to night-time
observing).  Centaurus~A~\citep{culgoora1} and M1~\citep{Braude1969}
affect the total power expected from the Global Sky
Model~\citep{gsm2008} at only the $\sim$10\% level.  Additionally, as
will be shown in \S\ref{s:obs}, preliminary \albatros\ data
demonstrate that the visibilities are only a few percent of the total
power amplitude, thus indicating that diffuse emission dominates even
the brightest sources over the large primary beam.

@@

As a qualitative illustration of the sky signal repeatability,
\autoref{Fig:albatros2_phase} shows the phases of the cross-spectra
binned in local sidereal time.  About 372 hours of data are averaged
into this plot, and no RFI excision has been performed.  A more
detailed analysis of signal repeatability will be presented in a
future paper, but the high signal-to-noise fringe pattern, which is
even visible slightly below $\sim10$~MHz, demonstrates the proof of
concept for building the expanded \albatros\ array.

>>>>

To assess the loss of sensitivity as the signal level increases, we
use \autoref{eqn:1bit_output} to derive the correlator efficiency
$\epsilon$, which is the ratio of measured signal-to-noise ratio (SNR)
to the ideal, infinite-precision SNR.  \autoref{fig:1bit_efficiency}
illustrates that the decrease in $\epsilon$ as a function of SNR is
relatively shallow.  When the SNR is near zero, $\epsilon$ approaches
$2/\pi \approx 0.64$, and if the signal level increases to two times
the noise level, $\epsilon$ decreases by only a factor of $\sim2$.
The final \albatros\ array will have baselines that are long relative
to the observing wavelength, and the correlated component of the
electric field (which sets the relevant signal level) is likely a
small fraction of the total received power.  This qualitative argument
can by made by considering antennas in the $uv$~plane in the flat-sky
approximation.  Any individual antenna receives all power present in
the $uv$~plane (plus system noise), but the correlated contribution
comes only from the portion of the $uv$~plane that is encompassed by
the primary beam at the antenna's $uv$-space coordinate.  Assuming
that the sky can be described as a Gaussian random field, the maximum
correlated signal between two antennas is determined by the aperture
filling factor of that pair.  For the case of dipole antennas, the
filling factor is the square of the wavelength divided by the baseline
length.  Therefore, in the long-baseline limit, the correlated signal
is small in comparison to the total power received, and the correlator
efficiency is unlikely to decrease significantly from the ideal
$2/\pi$ value.  This efficiency behavior remains true even when the
brightest radio sources are visible, since the primary beam
encompasses more than a steradian solid angle (the sun is irrelevant
since ionospheric effects restrict \albatros\ to night-time
observing).  Centaurus~A~\citep{culgoora1} and M1~\citep{Braude1969}
affect the total power expected from the Global Sky
Model (GSM)~\citep{gsm2008} at only the $\sim$10\% level.  Additionally, as
will be shown in \S\ref{s:obs}, preliminary \albatros\ data
demonstrate that the visibilities are only a few percent of the total
power amplitude, thus indicating that diffuse emission dominates even
the brightest sources over the large primary beam.

@@

We can estimate the instrumental gain stability by comparing the total
power between the different days shown in
\autoref{Fig:albatros2_waterfalls}.  We find that the RMS gain
fluctuations between 30 and 40~MHz (where the Galactic signal peaks in the
data) are less than 1\%.  Further, we find that the typical noise on
$\sim$3-second timescales is 0.04\% of the power.  The low noise suggests that we
can carry out relative calibration with sub-percent accuracy by using
the autospectra recorded by each autonomous station (absolute calibration
could be carried out by comparison to either \prizm\, which is
co-located and absolutely calibrated, or \textit{e.g.} the GSM).
As a qualitative illustration of the smaller-scale sky signal repeatability,
\autoref{Fig:albatros2_phase} shows the phases of the cross-spectra
binned in local sidereal time.  About 372 hours of data are averaged
into this plot, and no RFI excision has been performed.  A more
detailed analysis of signal repeatability will be presented in a
future paper, but the high signal-to-noise fringe pattern, which is
even visible slightly below $\sim10$~MHz, demonstrates the proof of
concept for building the expanded \albatros\ array.
